Bandwidth-Aware Reconfigurable Memory Design with Hybrid Memory Technologies

### 2011-04-06 trunk/paper.tex

Although the above hybrid memory system maintains the optimal provided bandwidth
curve over all the capacities, it does not garantee the best performance for
different applications with a variety of bandwidth requirements. The overall
provided bandwidth curve shown in Figure~\ref{fig:pbs-envelope} appears to be
monotonically decreasing. A specific application may have a bandwidth
requirement that can be mapped to a single point on the curve. To address this
issue, we dynamically reconfigure the memory space of each level corresponding
to the bandwidth requirement of each set of applications running on the system.
To be more aggressive, the reconfiguration is applied at the end of each given
time interval to accomodate different program phases of an application.

reconfiguration support: how reconfigurable (Li -> Li+1 (Li+0.5); hardware
modification to support reconfiguration; what to do with reconfigured capacity

decide size of each level si:

1) theoritical curves

2) bound:

- dbwi < pbwi -> si <= s1

- max throughput under power budget? -> si >= s2

- si <= hardware given max si

- si > si-1

>>>>

Although the above hybrid memory configuration maintains the optimal provided
bandwidth curve over all the capacities, it does not garantee the best
performance for different applications with a variety of bandwidth requirements.
The overall provided bandwidth curve shown in Figure~\ref{fig:pbs-envelope}
appears to be monotonically decreasing. A specific application may have a
bandwidth requirement that can be mapped to a single point on the curve. In
order to satisfy the bandwidth demand of different applications, we dynamically
reconfigure the memory space of each level corresponding to the bandwidth
requirement of each set of applications running on the system. To be more
aggressive, the reconfiguration is applied at the end of each given execution
time interval to accomodate different program phases of applications.

The major system parameter to be reconfigured is the capacity of each memory
level. At each time interval, we determine the upper bound of the capacity of
memory level-i ($s_{iu}$) by mapping the demand bandwidth (DBW) of a specific
application to the system's provided bandwidth curve, i.e.,
$s_{iu}=f^{-1}(\textrm{DBW})$ where $f(x)$ represents the provided bandwidth
curve of the on-chip memory system stored in the look-up-table in a reserved
non-volatile memory space. The demand bandwidth for each memory level is
generated using the prediction engine presented in section~\ref{sec:predict}. In
previous sections, we show that the memory with smaller capacity provides higher
bandwidth. Theoritically, the higher bandwidth provided by the memory system
leads to higher performance (in terms of throughput) and power consumption of
throughput computing applications~\cite{throughput-vs-bandwidth}. Consequently,
we define a lower bound of memory capacity as
$s_{il}=f^{-1}(\textrm{DBW}*(1+\sigma))$, where $\sigma$ is a pre-defined
threshold to make sure possible provided bandwidth increase without much power
overhead. The capacity of memory level-i ($s_i$) is therefore selected to
satisfy $s_{il}\le s_i\le s_{iu}$.

Our primary reconfigurable design exploits set associativity in conventional
cache organizations. Figure~\ref{fig:reconfig} depicts the block diagram for the
organization of our design. An n-way set associative cache consists of $n$ data
and tag arrays. We divide the each level of on-chip memory into partitions at
the granularity of the $k$-ways, where $k$ is determined by the number of
available capacity values for a specific memory technology on the provided
bandwidth curve. For example, if there are 8 available capacities for STT-RAM
(from 2MB to 1GB), we will determine the granularity by $k=n/8$.The desirability
of such configuration is to exploit the division of ways already present in a
conventional cache organization.

The index part of the input address is used to index all the
ways of the data and tag array. The tag part of the input address is sent to the
comparators of all the ways to determine if there is a match with any of the
tags read from the tag array. If there is a match, a hit is signaled on the
valid output line and data from the corresponding way of the data array is sent
onto the output data lines.

reconfiguration support: how reconfigurable (Li -> Li+1 (Li+0.5); hardware
modification to support reconfiguration.
